---
title: |
  \Large\bf Costs of primary healthcare visits and hospitalizations for scabies and bacterial skin infections in Fiji  \\
  Supplementary material

author:
  - name: Edifofon Akpan
    affil-id: 1, 2
  - name: Li Jun Thean
    affil-id: 3, 4
  - name: Maria Mow
    affil-id: 3  
  - name: Lucia Romani
    affil-id: 5
  - name: Joseph Kado
    affil-id: 6, 7
  - name: John Kaldor
    affil-id: 6
  - name: Andrew Steer
    affil-id: 3, 4
  - name: Natalie Carvalho
    affil-id: 1
    
affiliations: 
  - id: 1
    name: Melbourne School of Population and Global Health, The University of Melbourne, Australia 
  - id: 2
    name: Sheffield Centre for Health and Related Research, The University of Sheffield, United Kingdom
  - id: 3
    name: Murdoch Children's Research Institute, The Royal Children's Hospital, Australia
  - id: 4
    name: Department of Paediatrics, The University of Melbourne, Australia 
  - id: 5
    name: Kirby Institute, University of New South Wales, Kensington, Australia.
  - id: 6
    name: College of Medicine, Nursing and Health Sciences, Fiji National University, Suva, Fiji.
  - id: 7
    name: Wesfarmers Centre for Vaccines and Infectious Diseases, Telethon Kids Institute, Western Australia, Australia.
  - id: 8
    name: Ministry of Health and Medical Services, Suva, Fiji

crossref:
  title-delim: "."

number-sections: true

format:
  pdf:
    papersize: a4
    fontsize: "12pt"
    link-citations: true
    linestretch: 1.5
    geometry:
    - top=25.4mm
    - left=25.4mm
    - right=25.4mm
    - bottom=25.4mm
    template-partials: 
      - title.tex
    header-includes: |
      \usepackage[labelsep=period, labelfont=bf, textfont=normal]{caption}
      \usepackage[noblocks]{authblk}
      \renewcommand*{\Authsep}{, }
      \renewcommand*{\Authand}{, }
      \renewcommand*{\Authands}{, }
      \renewcommand\Authfont{\large} 
      \renewcommand\Affilfont{\normalsize}
  html: default
  
# mainfont: "Times New Roman"
# sansfont: "Times New Roman"
# monofont: "Consolas" 
# mathfont: "Cambria Math"

bibliography: references.bib

csl: ama.csl
---

```{r setup}
#| output: false
#| echo: false

library(tidyverse)
library(gtsummary)
library(flextable)
library(officer)

knitr::opts_chunk$set(fig.width = 12, fig.height = 10, fig.align = "center",
                      out.width = "100%", collapse = TRUE, fig.pos = "H", out.extra = "")

knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark=",")})

set_flextable_defaults(theme_fun = theme_booktabs, tabcolsep = 4, split = NULL, 
                       keep_with_next = NULL, border.width = 0.5)

ftTheme <- function(table){
  table <- table %>% 
    width(width = tblwidth) %>%
    align(i = NULL, j = NULL, align = "right",  part = "all") %>% 
    align(i = 1,    j = NULL, align = "center", part = "header") %>% 
    align(i = NULL, j = 1,    align = "left",   part = "all") %>% 
    fontsize(size = 11, part = "all") %>% 
    border_inner(border = fp_border(color="black", width = 0.5), part = "body")  %>% 
    border_outer(border = fp_border(color="black", width = 0.5), part = "body")  %>% 
    border_inner(border = fp_border(color="black", width = 0.5), part = "header") %>% 
    border_outer(border = fp_border(color="black", width = 0.5), part = "header")
  table
}

```

# Introduction {#sec-introduction .unnumbered}

The estimated global point prevalence of scabies reported in the 2016 Global Burden of Disease study was approximately 187 million and the annual incident case number was approximately 565 million. Scabies prevalence was 36.4% and impetigo prevalence was 23.4% in the Skin Health Intervention Fiji Trial (SHIFT) project conducted during the period 2012--2013 among the residents of the three islands of Fiji.[@romani2017] In 2017, The World Health Organization (WHO) recently acknowledged scabies as one of the neglected tropical diseases (NTDs) following recommendation from Strategic and Technical Advisory Group for Neglected Tropical Diseases.[@worldhealthorganization2017].

{{< pagebreak >}}

# Tables {#sec-tables .unnumbered}

```{=latex}
\renewcommand{\baselinestretch}{1}\normalsize
```
```{r}
#| echo: false
#| message: false
#| label: tbl-table1
#| tbl-cap: Characteristics of primary health care and hospitalized patients with scabies or skin and soft tissue infections in Northern Division, Fiji

table1   <- readRDS(file = "output/table1.rds")
tblwidth <- c(1.75, 1.00, 1.00, 1.00, 1.00)
table1   <- add_footer_lines(table1, value = as_paragraph(c("Percentages may not total 100 due to rounding. Raw numbers may not total sample size because of missing data in sex (75), age (197), and ethnicity (232 missing) variables. IQR, interquartile range; PHC, primary health care; NA, not applicable; SSTIs, skin and soft tissue infections.")))
ftTheme(table1)
```

{{< pagebreak >}}

```{r}
#| echo: false
#| message: false
#| label: tbl-table2
#| tbl-cap: Healthcare resource utilization and costs (in 2020 US dollars) of primary health care patients with scabies or skin and soft tissue infections.

stable2   <- readRDS(file = "output/stable2.rds")
tblwidth <- c(1.75, 1.00, 1.00, 1.00, 1.00)
stable2   <- add_footer_lines(stable2, value = as_paragraph(c("Percentages may not total 100 due to rounding. Raw numbers may not total sample size because of missing data in sex (75), age (197), and ethnicity (232 missing) variables. IQR, interquartile range; PHC, primary health care; NA, not applicable; SSTIs, skin and soft tissue infections.")))
ftTheme(stable2)
```

{{< pagebreak >}}

```{r}
#| echo: false
#| message: false
#| label: tbl-table3
#| tbl-cap: Healthcare resource utilization and costs (in 2020 US dollars) of hospitalized patients with scabies or skin and soft tissue infections.

stable3  <- readRDS(file = "output/stable3.rds")
tblwidth <- c(1.75, 1.00, 1.00, 1.00, 1.00)
stable3   <- add_footer_lines(stable3, value = as_paragraph(c("Percentages may not total 100 due to rounding. Raw numbers may not total sample size because of missing data in sex (75), age (197), and ethnicity (232 missing) variables. IQR, interquartile range; PHC, primary health care; NA, not applicable; SSTIs, skin and soft tissue infections.")))
ftTheme(stable3)

```

{{< pagebreak >}}

# References {#sec-references .unnumbered}

::: {#refs}
:::
